A Database Management System for a Geospatially-Enabled Interactive Game

### 2018-10-27 paper/MSDS7330DBMSforGeospatiallyEnabledGame.tex
Minor changes in formatting
Minor Edits
Added to Database Design Section

\usepackage{float}
\usepackage{graphicx}
\usepackage{hyperref}

@@

\section{Game rules and design considerations}
\subsection{Problem statement}
As mentioned, the game will consist of players exploring the real world in search of tokens. Collecting tokens raises the player’s score which will in turn reveal a segment of a story once the player’s score reaches a certain threshold. The tokens belong to different categories that correspond to different aspects of the story that may be revealed. As such, this game relies heavily on the search-and-find mechanic as well as desires for collection, completion, and closure of narrative for player engagement \cite{game-methodology, location-based-games}. 
The story is a science fiction story that places the player in the role of an explorer. The player finds himself exploring ruins on another world that once housed an advanced civilization. The player must then find tokens that will divulge information about this lost civilization including clues as to what caused the civilization to disappear. Categories consist of Culture, Politics, Technology, and Economy and correspond to the nature of the clues the player discovers. 
Given the rules, mechanics, and the location-based nature of the game, some data and database design considerations are:

@@

\begin{itemize}
  \item Client-side application. We decided to support 2 major platforms that got widespread adoption: Android and iOS. For the proof-of-concept we focused on iOS. 
  \item Backend services, or middle layer. This component performs server-side operations, enforces gameplay rules and handles requests to our database. 
  \item Geospatially enabled database. Here all information regarding our players, tokens and corresponding states is being kept. 
\end{itemize}

@@

\begin{itemize}
\item REST over HTTP. 
\item SQL 
\end{itemize}

@@

\section{Database design}
TBD
\section{Conclusion}
TBD

>>>>

\usepackage{float}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{enumitem}

@@

\section{Game rules and design considerations}
\subsection{Problem statement}
As mentioned, the game will consist of players exploring the real world in search of tokens. Collecting tokens raises the player’s score which will in turn reveal a segment of a story once the player’s score reaches a certain threshold. The tokens belong to different categories that correspond to different scores that must be earned and different aspects of the story that may be revealed. As such, this game relies heavily on the search-and-find mechanic as well as desires for collection, completion, and closure of narrative for player engagement \cite{game-methodology, location-based-games}.

The story is a science fiction story that places the player in the role of an explorer. The player finds himself exploring ruins on another world that once housed an advanced civilization. The player must then find tokens that will divulge information about this lost civilization including clues as to what caused the civilization to disappear. Categories consist of Culture, Politics, Technology, and Economy and correspond to the nature of the clues the player discovers.

Given the rules, mechanics, and the location-based nature of the game, some data and database design considerations are:

@@

\begin{enumerate}
  \item Client-side application. We decided to support 2 major platforms that got widespread adoption: Android and iOS. For the proof-of-concept we focused on iOS. 
  \item Backend services, or middle layer. This component performs server-side operations, enforces gameplay rules and handles requests to our database. 
  \item Geospatially enabled database. Here all information regarding our players, tokens and corresponding states is being kept. 
\end{enumerate}

@@

\begin{enumerate}[label=\Alph*]
\item REST over HTTP. 
\item SQL 
\end{enumerate}

@@

\section{Database design}

The important entities that the database must be store are player data, story-segments, token-category properties and token locations. As developers, we also need to know a player’s score in order to know what story segments the player has earned. Similarly, we need to know what tokens have already been collected in order to not allow the player to collect them again.

One major consideration in choice of database schema is to what extent the game may be housed on the player’s phone and what may be streamed from the central database. The former makes data integrity more feasible while the latter makes the storage of the player’s progress more feasible. For example, knowing what tokens the player has already collected entails storing information that depends on both the player base and all of the tokens’ geospatial locations. Hence, it merits its own table the size of the number of players times the number of token locations.  As the player base increases and the game is expanded, this table will grow very rapidly in size. However, if the player holds the token locations portion of the database on his or her phone, then the information may simply be stored as an indicator on the player’s downloaded schema.

\section{Conclusion}
TBD
